---
title: "R Notebook"
output: html_notebook
---


```{r, echo = false, include=FALSE}
library(ggplot2)
library(dplyr)
```


## Data exploration
Starting with the zenodo download from https://zenodo.org/record/4953762#.YYQmzJso_s2, we explore the files first. We have 6 sites either in Portugal or in Denmark. 

```{r}
data_dir <- "../data/raw-data/"

read.csv(paste0(data_dir, "01_Sites_coord.csv"))
# UTM Coordinates need clean-up, bc all information in one cell TODO

read.csv(paste0(data_dir, "02_Polygons_coord.csv"))
# Coordinates need clean-up, bc not in tidy format, center coords given

# Pesticide applications on one site in Denmark
read.csv(paste0(data_dir, "Table_I_Pesticide_application.csv"))

# Resources
# =========
reso <- read.csv(paste0(data_dir, "Table_II_Resource_providing_unit_coord.csv"))
# observations per site

# Observations per polygon
x <- reso %>% 
  group_by(siteNo, polygonId) %>%
  summarise(n = n())
hist(x$n,  breaks = 300)

# only keep sites with at least 6 observations? 

# observations per month & year

x <- reso %>% 
  group_by(siteNo, sampY, sampM) %>%
  summarize(n = n())

ggplot(data=x, aes(x = factor(sampM), y = n)) + 
  facet_grid(rows = vars(sampY)) + 
  geom_boxplot()

```





```{r}

table(reso$siteNo)
# observations per polygon 
hist(sort(table(reso$polygonId), decreasing = TRUE))

# observations per month 
table(reso$sampM)
# resource availability in classes, 1 = low, 3 = high
table(reso$abundanceClass)
names(reso)
library(dplyr)
reso2 <- reso %>% 
  group_by(resId, sampY, sampM, sampD, polygonId) %>%
  summarize(abundanceClass = abundanceClass)  

table(reso$siteNo)
# siteNo not unique

read.csv(paste0(data_dir, "Table_III_Hives.csv"))
# gives hive IDs

read.csv(paste0(data_dir, "Table_IV_Colony_management.csv"))

```

## Colony performance

```{r}

perf <- read.csv(paste0(data_dir, "Table_V_Colony_inspection.csv"))


# Colony weight

# n observations per month & year

x <- perf %>% 
  group_by(siteNo, sampY, sampM) %>%
  filter(paramText == "Number of adult bees") %>%
  summarize(n_observations = n())

ggplot(data=x, aes(x = factor(sampM), y = n_observations)) + 
  facet_grid(rows = vars(sampY)) + 
  geom_boxplot()

# TODO: function to calculate measurement day (day of year) from sampM, sampD 

```



```{r}


table(perf$resId) # MSP thoughts: it looks like resID is unique per observation and contains no site/hive information? 


table(perf$siteNo)
table(perf$hiveNo)

# How many observations per hive?   
perf %>% group_by(resId, sampY, sampM, sampD, paramText) %>%
  summarize(n_observations = n())


# Adult bees
# ==========

table(perf$paramText)
names(perf) 

perf_adult_bees <- perf %>% 
  group_by(resId, sampY, sampM, sampD, resVal) %>%
  filter(paramText == "Number of adult bees") %>%
  summarize(adult_bees = resVal)





ggplot(data = perf_adult_bees, aes(x = sampM, y = adult_bees))+ 
  geom_point()
# Colony size seems too low? (< 8000 for all colonies) 

perf_col_weight <- perf %>% 
  group_by(resId, sampY, sampM, sampD, resVal) %>%
  filter(paramText == "Colony weight") %>%
  summarize(colony_weight = resVal,
            n_records = n())

count(perf_col_weight$resVal)

ggplot(data = perf_col_weight, aes(x = factor(sampM), y = colony_weight))+ 
  geom_point(col = "grey")+
  geom_boxplot()+
  ylim(0, max(perf_col_weight$colony_weight))

# How many records per colony? 




read.csv(paste0(data_dir, "Table_VI_SSD2.csv"))

read.csv(paste0(data_dir, "Table_VII_Observation_colony.csv"))

# subset 1:
# One location with monthly records of resources, enough hive records, etc 

# can we link resources and performance? 
sum(reso$resId  %in% perf$resId) #  no single match between IDs, 

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
